%==================================================================================
%==================================================================================
% Document		:		Annexe: Liste des acronymes
%
% Auteur		: 		Abdelkrime ARIES
% Encadreur		:		Dr. Omar NOUALI
% Co-encadreur	:		Mme. Houda OUFAIDA
% Établissement	:		ESI (Ecole Nationale Supérieure d'Informatique; ex. INI) 
% Adresse		:		Oued Smar, Alger, Algérie 
% Année			:		2012/2013
% Grade			:		Magister
% discipline 	:		Informatique 
% Spécialité	:		IRM (Informatique Répartie et Mobile)
% Titre			:		Résumé automatique de textes
%
%==================================================================================
%==================================================================================

%==========================L'entete de chapitre====================================
%==================================================================================
 \ifx\wholebook\relax\else
  	\documentclass[a4paper,12pt,oneside]{../use/ESIthesis}
  	
  	\input{../use/formatAndDefs}
  	 	
  	 	\setracine{../}
  	 	\graphicspath{{.}{../fig/}}

  	 	\begin{document}

  	 	\selectlanguage {francais}

  	 	\mainmatter
  	 	\appendix
  \fi
%==================================================================================
\chapter{Publications personnelles}
\noindent
{\huge Using clustering and a modified classification algorithm for automatic text summarization}\\[20px]
%Abdelkrime Aries ; Houda Oufaida ; Omar Nouali\\
%Ecole Nationale Supérieue d'Informatique (Algeria)\\
%Ctr. de recherche sur l'Information Scientifique et Technique (Algeria)\\
\textbf{Abdelkrime Aries, Houda Oufaida}\\
Ecole Nationale Supérieue d'Informatique (Algeria)\\[3mm]
%
\textbf{Omar Nouali}\\
Ctr. de recherche sur l'Information Scientifique et Technique (Algeria)\\
%
\kbox{\small 
\textbf{Proc. SPIE Volume 8658}\\[5px]
\hspace*{10mm}
Document Recognition and Retrieval XX, 865811\\
\hspace*{10mm}
Richard Zanibbi; Bertrand Coüasnon\\
\hspace*{10mm}
Burlingame, California, USA | February 03, 2013\\
\hspace*{10mm}
\url{http://dx.doi.org/10.1117/12.2004001}\\
}\\[5mm]
%
{\Large abstract}\\
In this paper we describe a modified classification method destined for extractive summarization purpose. 
The classification in this method doesn’t need a learning corpus; it uses the input text to do that. 
First, we cluster the document sentences to exploit the diversity of topics, then we use a learning algorithm (here we used Naive Bayes) on each cluster considering it as a class. 
After obtaining the classification model, we calculate the score of a sentence in each class, using a scoring model derived from classification algorithm. 
These scores are used, then, to reorder the sentences and extract the first ones as the output summary. 
We conducted some experiments using a corpus of scientific papers, and we have compared our results to another summarization system called UNIS. 
Also, we experiment the impact of clustering threshold tuning, on the resulted summary, as well as the impact of adding more features to the classifier. 
We found that this method is interesting, and gives good performance, and the addition of new features (which is simple using this method) can improve summary’s accuracy.

%========================Le pied de chapitre=======================================
%==================================================================================
\ifx\wholebook\relax\else
 \cleardoublepage
 \bibliographystyle{../use/ESIbib}
 \bibliography{../bib/RA}
 \end{document}
\fi
%==================================================================================